---
title: "p8105_hw3_lc3807.Rmd"
author: "Linshen Cai"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}

# load necessay package
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)
```


## Problem 1 

## Load data set

```{r}

library(p8105.datasets)
data("instacart")

instacart = 
  instacart |> 
  as_tibble()

```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.


## Summarizing the number of items ordered from aisle
```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

Above is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.


## Plot the scatter plot
```{r}

instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

Above is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items. 

## Show the three most popular items
```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

The table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

## The mean hour of the day
```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```

Finally, this is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.


## Problem 2

## Load the data and do some data cleaning
```{r}
library(p8105.datasets)
data("brfss_smart2010")

# data cleaning
brfss_smart2010 = 
  brfss_smart2010 |> 
  as_tibble() |> 
  janitor::clean_names()

brfss_smart2010 = brfss_smart2010 |> 
  filter(topic == "Overall Health",
         response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")
         ) |> 
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))
```
This dataset contains `r nrow(brfss_smart2010)` rows and `r ncol(brfss_smart2010)` columns, with each row representing the number of response about their overall health from "Poor" to "Excellent" in different location of all states every year.

```{r}

# count states with 7 or more location oberved in 2020
state_counts_2002 = brfss_smart2010 |> 
  filter(year == 2002) |> 
  group_by(locationabbr) |> 
  summarize(NumObservations = n())
states_7_or_more_2002 = state_counts_2002 |> 
  filter(NumObservations >= 7)
# count states with 7 or more location oberved in 2010
state_counts_2010 = brfss_smart2010 |> 
  filter(year == 2010) |> 
  group_by(locationabbr) |> 
  summarize(NumObservations = n())
states_7_or_more_2010 = state_counts_2010 |> 
  filter(NumObservations >= 7)
 
```

In 2002, there are `r nrow(states_7_or_more_2002)` states were observed at 7 or more locations. They are `r states_7_or_more_2002 |> pull(locationabbr)`.
In 2010, there are `r nrow(states_7_or_more_2010)` states were observed at 7 or more locations. They are `r states_7_or_more_2010 |> pull(locationabbr)`.

```{r}
# Filter the data for "Excellent" responses
response_excellent = brfss_smart2010 |> 
  filter(response == "Excellent") 

state_response_excellent = response_excellent |> 
  group_by(year,locationabbr) |> 
  summarize(mean_data_value = mean(data_value, na.rm = TRUE))

state_response_excellent |> 
  ggplot(aes(x = year, y = mean_data_value, color = locationabbr)) +
  geom_line() +
  labs(title = "Averages Data_value for Each State",
      x = "Years",
      y = "Averages data_value")
  
```

Above is plot showing averages data_value across locations for each state across years.

```{r}
state_counts_2006_2010 = brfss_smart2010 |> 
  filter(year == 2006 | year == 2010, locationabbr == "NY")

state_counts_2006_2010 |> 
  ggplot(aes(x = response, y = data_value, fill = year )) +
  geom_boxplot()+ 
  facet_wrap(~ year) +
  labs(title = "Distribution of Data Value by Response for NY State in 2006 and 2010",
       x = "Response",
       y = "Data Value") +
  theme_minimal()
  
  
  
```

